---
title: "DataFest Workspace"
author: "Elizabeth"
date: "April 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages = FALSE, fig.align = "center", cache = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("."))
options(stringsAsFactors = FALSE)
```

```{r read_data}
setwd("C:/Users/eabert/Documents/Programming/analytics_projects/DataFest 2019/datatastic_2019/datafest-osu-2019/data")
games <- read.csv(file = "games.csv")
gps <- read.csv(file = "gps.csv")
# rpe <- read.csv(file = "rpe.csv")
wellness <- read.csv(file = "wellness.csv")
wellnessMessy <- wellness
```

## Sleep EDA

```{r start_subsetting}
summary(wellness[7:10])
p1_bt <- wellness$BedTime[wellness$PlayerID == 1]
p1_bt <- subset(wellness, wellness$PlayerID == 1)
```

```{r sleep_eda}
library(dplyr)

tapply(wellness$BedTime, wellness$PlayerID, summary) # successful summary, failed to hist

# successful summary
wellness %>% 
  group_by(PlayerID) %>% 
  summarize(mean = mean(SleepHours), sd = sd(SleepHours)) 

# successful plot
wellness %>% 
  group_by(PlayerID) %>% 
  summarize(mean=mean(SleepHours)) %>% 
  plot()

# successful boxplot
## Lots of variance/outliers
wellness %>%
  boxplot(SleepHours ~ PlayerID, data = .)

library(ggplot2)
ggplot(wellness, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
```

## Data cleaning - Reporting Mistakes

### Military Time Mistake

#### Length Cleaning:
Subtract 12 hours to account for military time mistake from sleep times over 19

```{r shorten_long}
# TODO Look at these rows and see if they needed to be cleaned like this

# some people slept way too much
cutoff <- 19
# base R
wellness$SleepHours[wellness$SleepHours > cutoff] = wellness$SleepHours[wellness$SleepHours > cutoff] - 12
# dplyr alternative
wellness %>%
  mutate(SleepHours = 
           case_when(
             SleepHours > cutoff ~ SleepHours - 12,
             TRUE ~ SleepHours
           ))
```

#### Bed/Wake Time Cleaning

##### Found Errors
```{r find_errors}
# because of the distribution of bedtimes, assume they meant to use 24 hour time and no one went to sleep mid-morning
hist(wellness$BedTimeHour)

# reporting error verified - look at people asleep before noon (military v standard time)
require(lubridate)
wellness %>%
  filter(hour(BedTime) <= 12)

# time as an integer for easy (imprecise!) comparison, instead of lubridate
wellness$BedTimeHour <- as.integer(substring(wellness$BedTime, 0, 2))
wellness$WakeTimeHour <- as.integer(substring(wellness$WakeTime, 0, 2))

# predict wake time
wellness$PredWakeTime <- (wellness$BedTimeHour + wellness$SleepHours) %% 24
# check
wellness$change12hr <- (abs(trunc(wellness$PredWakeTime) - wellness$WakeTimeHour) > 2)
# verify predicted wake time - changed to filter (better)
wellness %>%
  # filter(BedTimeHour > 4 & BedTimeHour < 12 & wellness$change12hr == TRUE) %>% 
  filter(wellness$change12hr == TRUE) %>%
  # select(c(PlayerID, BedTimeHour, SleepHours, PredWakeTime, WakeTimeHour, change12hr)) %>% 
  View()
```

##### Naiive Clean Errors by Dropping

```{r drop_time}
wellnessClean <- wellness %>%
  filter(change12hr == FALSE)
```

##### Clean Errors
```{r clean_time}
# TODO NEXT FIND OUT HOW TO ADD 12 HOURS

# TODO - sub 12 from wake for __player 17 on 2017-10-15__ and __player 17 on 2018-01-02__
wellness %>%
  mutate(WakeTime = 
           case_when(
             PlayerID == 17 & Date == "2017-10-15" ~ WakeTime - 12,
             PlayerID == 17 & Date == "2018-01-02" ~ WakeTime - 12,
             TRUE ~ WakeTime
           ))

# TODO rerun flag code here

# TODO - add 12 hrs to sleep for __flag__
wellness %>%
  mutate(BedTime = 
           case_when(
             change12hr == TRUE ~ BedTime + 12,
             TRUE ~ BedTime
           ))
```

### Data Cleaning Result Impact

```{r comparison}
# From sleep_eda:
wellnessMessy %>%
  boxplot(SleepHours ~ PlayerID, data = .)
wellnessClean %>%
  boxplot(SleepHours ~ PlayerID, data = .)

ggplot(wellnessMessy, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
ggplot(wellnessClean, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
# TODO fix axes
```

```{r wellness_export}
# Shared via Slack
wellnessClean %>%
  write.csv(file = "wellness_clean.csv")
```

## Wellness/Fatigue EDA
Correlation between wellness metrics, paying special attention to corr to fatigue

```{r}
library(GGally)
full_list <- c("Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality", "MonitoringScore", "Pain", "Illness", "Menstruation", "Nutrition", "NutritionAdjustment", "USGMeasurement", "USG")
short_list <- c("Fatigue", "Soreness", "Desire")
wellnessClean %>%
  select(full_list) %>% 
  ggpairs(upper = list(continuous = wrap("cor", size = 9)))

cor_numeric <- c("Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality") #, "MonitoringScore")
wellnessClean %>%
  select(cor_numeric) %>% 
  cor() %>%
  View()

round(cor(select(wellnessClean, cor_numeric))[3], 4) # 1) correlation with desire
round(cor(select(wellnessClean, cor_numeric))[6], 4) # 2) correlation with sleep quality
round(cor(select(wellnessClean, cor_numeric))[2], 4) # 3) correlation with soreness

```

## Fatigue to RPE

```{r}
summary(rpe)
table(rpe$Training) # 7622 yes
table(rpe$SessionType)
length(rpe$SessionType) - 463 # = 8397 sessions > 7622 training days, therefore multiple training sessions/day

rpe %>%
  filter(SessionType == "Game") %>%
  group_by(Date, PlayerID) %>%
  summarize(game_train = sum(which(Training == "Yes"))) %>%
  pull(game_train) %>%
  hist()
# more complicated - what other types of training did they have on days they had games?

rpe %>%
  filter(SessionType == "Game") %>%
  group_by(Date, PlayerID) %>%
  summarize(game_train = sum(which(Training == "Yes"))) %>% # everyone had lots of sessions even on their game days
  filter(game_train == 0)
# every day they had a game they also trained

rpe %>% 
  filter(PlayerID == "2" & Date == "2017-10-28")
# 3 games and 1 skills training

rpe %>%
  filter(Training == "No") %>%
  group_by(Date, PlayerID) %>% # this line doesn't do anything; only one observation per player per off day
  nrow()
# 1238 off days

rpe %>%
  filter(Training == "No" & !is.na(SessionType))
# player 3 didn't train but did do conditioning on 2017-11-30

# Game Day Only Data:
game_days <- rpe %>%
  filter(SessionType == "Game") %>%
  pull(Date) %>%
  unique()
rpe %>%
  filter(Date %in% game_days) %>%
  View()
# They do lots of other things around when they're playing

# Game day workouts per player:
gameday_rpe <- rpe %>%
  filter(Date %in% game_days) %>%
  group_by(Date, PlayerID) %>%
  mutate(Sessions = paste0(SessionType, collapse = ", ")) %>%
  group_by(Date, PlayerID) %>%
  summarize(total_duration = sum(Duration), total_RPE = sum(RPE), total_SessionLoad = sum(SessionLoad), Sessions = first(Sessions), DailyLoad = first(DailyLoad), AcuteLoad = first(AcuteLoad), ChronicLoad = first(ChronicLoad), AcuteChronicRatio = first(AcuteChronicRatio), ObjectiveRating = first(ObjectiveRating), FocusRating = first(FocusRating), BestOutOfMyself = first(BestOutOfMyself), Training = first(Training)) %>% # first() to save var
  filter(Training == "Yes") 
```

# Day 2

## Connecting meaningful things to RPE

```{r}
# Using gameday_rpe combine in wellness for those players and games
well_rpe <- merge(gameday_rpe, wellnessClean, by = c("Date", "PlayerID"), all = TRUE)
# drop NA rows
well_rpe <- well_rpe %>%
  filter(!is.na(total_duration))

# tried to order bedtimes as numbers - failed
well_rpe$BedTime <- as.numeric(well_rpe$BedTimeHour)
well_rpe %>%
  mutate(BedTimeHour = case_when(BedTimeHour < 12 ~ BedTimeHour + 24,
                                 TRUE ~ BedTimeHour)) %>%
  ggplot(aes(x = BedTimeHour, y = total_RPE)) +
  geom_point()
# clumsily added back to well_rpe - success in plotting sleep over time
well_rpe_late <- well_rpe %>%
  filter(BedTimeHour < 12) %>%
  mutate(BedTimeHour = BedTimeHour + 24)
well_rpe_early <- well_rpe %>%
  filter(BedTimeHour >= 12) 
well_rpe_new <- rbind(well_rpe_late, well_rpe_early)

ggplot(well_rpe_new, aes(x = BedTimeHour, y = total_RPE)) +
  geom_point()
cor_numeric <- c("total_duration", "ObjectiveRating", "FocusRating", "Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality", "MonitoringScore", "USG", "TrainingReadiness", "BedTimeHour", "WakeTimeHour")
# total_duration - work athletes longer, they get more exhausted
# ObjectiveRating - they can accomplish their objective without getting too much more exhausted (low slope)
# FocusRating - "" ""
# Fatigue - their workout isn't going to exhaust them more because they woke up fatigued
# Soreness - they can power through a workout while being sore
# Desire & Irritability & SleepHours & SleepQuality & dehydration no impact too
# TrainingReadiness - more ready to train, more exertion
# SleepTime

well_rpe_new %>%
  select(c("total_duration", "total_RPE", "ObjectiveRating", "FocusRating", "Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality", "MonitoringScore", "USG", "BedTimeHour", "WakeTimeHour")) %>% 
  cor() %>%
  View()
```

## Merge wellness rpe with games

```{r}
games$PointRatio <- games$TeamPoints / (games$TeamPointsAllowed)
well_rpe_game <- merge(well_rpe, games, by = c("Date"), all = TRUE)
well_rpe_game <- well_rpe_game %>%
  filter(!is.na(Outcome))

# % of time players we have data on win their games
nrow(well_rpe_game[well_rpe_game$PointRatio > 1,]) / nrow(well_rpe_game)

well_rpe_game_high <- well_rpe_game %>%
  filter(PointRatio == Inf) %>%
  mutate(PointRatio = 6.25)
well_rpe_game_other <- well_rpe_game %>%
  filter(PointRatio != Inf)
well_rpe_game_new <- rbind(well_rpe_game_high, well_rpe_game_other)

ggplot(well_rpe_game_new, aes(x = total_RPE, y = PointRatio)) +
  geom_point() +
  geom_abline()

well_rpe_game_new %>%
  select(c("PointRatio", "total_duration", "total_RPE", "ObjectiveRating", "FocusRating", "Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality", "MonitoringScore", "USG", "BedTimeHour", "WakeTimeHour")) %>%
  cor(use = "complete.obs") %>%
  View()
```

# TODO Go to Dave with 3 step thoughts wellness->RPE

## GPS Analysis
Data from Cole & Edwin - unexlplored/unvisualized

```{r}
setwd("C:/Users/eabert/Documents/Programming/analytics_projects/DataFest 2019/datatastic_2019/datafest-osu-2019/data")
distance <- read.csv(file = "distance.csv")
playtime <- read.csv(file = "playtime.csv")

# Summarize them
```

# TODO Later: make rolling sleep consistency metric (sd sleep time over 3 days? 1 week?)
