---
title: "DataFest Workspace"
author: "Elizabeth"
date: "April 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages = FALSE, fig.align = "center", cache = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("."))
options(stringsAsFactors = FALSE)
```

```{r read_data}
setwd("C:/Users/eabert/Documents/Programming/analytics_projects/DataFest 2019/datatastic_2019/datafest-osu-2019/data")
games <- read.csv(file = "games.csv")
gps <- read.csv(file = "gps.csv")
# rpe <- read.csv(file = "rpe.csv")
wellness <- read.csv(file = "wellness.csv")
wellnessMessy <- wellness
```

## Sleep EDA

```{r start_subsetting}
summary(wellness[7:10])
p1_bt <- wellness$BedTime[wellness$PlayerID == 1]
p1_bt <- subset(wellness, wellness$PlayerID == 1)
```

```{r sleep_eda}
library(dplyr)

tapply(wellness$BedTime, wellness$PlayerID, summary) # successful summary, failed to hist

# successful summary
wellness %>% 
  group_by(PlayerID) %>% 
  summarize(mean = mean(SleepHours), sd = sd(SleepHours)) 

# successful plot
wellness %>% 
  group_by(PlayerID) %>% 
  summarize(mean=mean(SleepHours)) %>% 
  plot()

# successful boxplot
## Lots of variance/outliers
wellness %>%
  boxplot(SleepHours ~ PlayerID, data = .)

library(ggplot2)
ggplot(wellness, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
```

## Data cleaning - Reporting Mistakes

### Military Time Mistake

#### Length Cleaning:
Subtract 12 hours to account for military time mistake from sleep times over 19

```{r shorten_long}
# TODO Look at these rows and see if they needed to be cleaned like this

# some people slept way too much
cutoff <- 19
# base R
wellness$SleepHours[wellness$SleepHours > cutoff] = wellness$SleepHours[wellness$SleepHours > cutoff] - 12
# dplyr alternative
wellness %>%
  mutate(SleepHours = 
           case_when(
             SleepHours > cutoff ~ SleepHours - 12,
             TRUE ~ SleepHours
           ))
```

#### Bed/Wake Time Cleaning

##### Found Errors
```{r find_errors}
# because of the distribution of bedtimes, assume they meant to use 24 hour time and no one went to sleep mid-morning
hist(wellness$BedTimeHour)

# reporting error verified - look at people asleep before noon (military v standard time)
require(lubridate)
wellness %>%
  filter(hour(BedTime) <= 12)

# time as an integer for easy (imprecise!) comparison, instead of lubridate
wellness$BedTimeHour <- as.integer(substring(wellness$BedTime, 0, 2))
wellness$WakeTimeHour <- as.integer(substring(wellness$WakeTime, 0, 2))

# predict wake time
wellness$PredWakeTime <- (wellness$BedTimeHour + wellness$SleepHours) %% 24
# check
wellness$change12hr <- (abs(trunc(wellness$PredWakeTime) - wellness$WakeTimeHour) > 2)
# verify predicted wake time - changed to filter (better)
wellness %>%
  # filter(BedTimeHour > 4 & BedTimeHour < 12 & wellness$change12hr == TRUE) %>% 
  filter(wellness$change12hr == TRUE) %>%
  # select(c(PlayerID, BedTimeHour, SleepHours, PredWakeTime, WakeTimeHour, change12hr)) %>% 
  View()
```

##### Naiive Clean Errors by Dropping

```{r drop_time}
wellnessClean <- wellness %>%
  filter(change12hr == FALSE)
```

##### Clean Errors
```{r clean_time}
# TODO NEXT FIND OUT HOW TO ADD 12 HOURS

# TODO - sub 12 from wake for __player 17 on 2017-10-15__ and __player 17 on 2018-01-02__
wellness %>%
  mutate(WakeTime = 
           case_when(
             PlayerID == 17 & Date == "2017-10-15" ~ WakeTime - 12,
             PlayerID == 17 & Date == "2018-01-02" ~ WakeTime - 12,
             TRUE ~ WakeTime
           ))

# TODO rerun flag code here

# TODO - add 12 hrs to sleep for __flag__
wellness %>%
  mutate(BedTime = 
           case_when(
             change12hr == TRUE ~ BedTime + 12,
             TRUE ~ BedTime
           ))
```

### Data Cleaning Result Impact

```{r comparison}
# From sleep_eda:
wellnessMessy %>%
  boxplot(SleepHours ~ PlayerID, data = .)
wellnessClean %>%
  boxplot(SleepHours ~ PlayerID, data = .)

ggplot(wellnessMessy, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
ggplot(wellnessClean, aes(SleepHours, SleepQuality)) +
  geom_point(alpha = 0.05)
# TODO fix axes
```

```{r wellness_export}
# Shared via Slack
wellnessClean %>%
  write.csv(file = "wellness_clean.csv")
```

## Wellness/Fatigue EDA
Correlation between wellness metrics, paying special attention to corr to fatigue

```{r}
library(GGally)
full_list <- c("Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality", "MonitoringScore", "Pain", "Illness", "Menstruation", "Nutrition", "NutritionAdjustment", "USGMeasurement", "USG")
short_list <- c("Fatigue", "Soreness", "Desire")
wellnessClean %>%
  select(full_list) %>% 
  ggpairs(upper = list(continuous = wrap("cor", size = 9)))

cor_numeric <- c("Fatigue", "Soreness", "Desire", "Irritability", "SleepHours", "SleepQuality") #, "MonitoringScore")
wellnessClean %>%
  select(cor_numeric) %>% 
  cor() %>%
  View()

round(cor(select(wellnessClean, cor_numeric))[3], 4) # 1) correlation with desire
round(cor(select(wellnessClean, cor_numeric))[6], 4) # 2) correlation with sleep quality
round(cor(select(wellnessClean, cor_numeric))[2], 4) # 3) correlation with soreness

```


# TODO Next: get weather data from Cole

# TODO Later: make rolling sleep consistency metric @ Cole (sd sleep time over 3 days? 1 week?)
